---
title: "convert"
author: "Robert J.D. Reid, Ph.D."
date: "2022-11-20"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

### Converting files from Volocity for use in R/Rstudio

Volocity 5.5 is an old old mac-based app for microscopy that uses
the legacy character code (Mac OS Roman = iso-8859-1) and legacy
Mac OS carriage returns ('\r').
These need to be converted to UTF-8 and unix line feeds ('\n').
Most efficient way to do this is with the system commands `iconv`
and `tr`.
As such, this is platform-specific - works on Macs & perhaps unix/
linux, but not on windows.


read file names and convert

Helper function defined first to call out to system
```{r}
# helper function calls out to system to convert mac CR into 'NIX newlines using `tr`
# if the original file has windows LFs this will result in 2 newline characters & blank lines
# so there is a pipe into a second `tr` that will remove blank lines with a "squeeze" switch

convert.breaks <- function(fls,dr_rd,dr_wr) {
  files_read <- file.path(dr_rd,fls) %>% str_replace_all(' ','\\\\ ')
  files_write <- file.path(dr_wr,fls) %>% str_replace_all(' ','\\\\ ')
  
  purrr::map2(files_read,files_write,function(x,y){
    cmd_string <- paste0("iconv -f iso-8859-1 -t utf-8 ",
                         x,
                         " | tr '\r' '\n' | tr -s '\n' > ",
                         y)
    system(cmd_string)
  })
}
```

Then logic to check folders

turn this into a function to provide folder names
```{r}
# add in checks for file existence, directory existence, directory create
# and overwrite warning
# turn this into a script that is sourced from the processing file
dir_read <- "c151-180"
files_source <- list.files(dir_read, pattern = "\\.csv")
dir_write <- "c151-180_conv"
if(dir.exists(dir_write)) {
  files_target <- list.files(dir_write, pattern = "\\.csv")
  if(any(files_source %in% files_target)) {
    stop ("files already in target directory")
  } else {
    convert.breaks(files_source,dir_read,dir_write)
    }
} else { # target dir doesn't exist so make it
  dir.create(dir_write)
  convert.breaks(files_source,dir_read,dir_write)
}
```


