---
title: "Micro_data_process"
author: "Robert Reid"
date: "2022-11-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(dendextend)
library(viridisLite)
knitr::opts_chunk$set(echo = TRUE)
```

### Notes on dataset

So Jung uses Volocity to ID fluorescent foci.
An object ID function is used to mark RFP and YFP.
A local maxima 'spot ID' function is used to ID CFP foci and object ID is used to find the whole nucleus

Note: The output from Volocity is using 'classic MAC OS' line breaks and Western MAC OS character set.

A convert-files script can be run from the `convert-files.Rmd` document
in this same folder.
Directory with converted files is `data_conv`

```{r}
dir <- 'data-conv'
```


### Read in 4 separate data sources across multiple files and combine

```{r read_data echo=FALSE,message=FALSE,warning=FALSE}
YFP_files <- list.files(dir, pattern = "YFP-foci", full.names = TRUE)


YFP <- YFP_files %>% map_dfr(~read_csv(.)) %>%
  mutate (field = str_remove(`Item Name`," \\(registration corrected\\)")) %>%
  separate(field, into = c("sample","field"), sep = " ") %>%
  replace_na(list(field="1")) %>%
  mutate(xfp_id = paste("yfp",str_remove(Name,"Object "))) %>%
  select(
    sample,
    #time,
    field,
    xfp_id,
    cenX = `Centroid X (µm)`,
    cenY = `Centroid Y (µm)`,
    cenZ = `Centroid Z (µm)`,
    vol = `Volume (µm3)`
  )

RFP_files <- list.files(dir, pattern = "RFP-foci", full.names = TRUE)
RFP <- RFP_files %>% map_dfr(~read_csv(.)) %>%
  mutate (field = str_remove(`Item Name`," \\(registration corrected\\)")) %>%
  separate(field, into = c("sample","field"), sep = " ") %>%
  replace_na(list(field="1")) %>%
  mutate(xfp_id = paste("rfp",str_remove(Name,"Object "))) %>%
  select(
    sample,
    #time,
    field,
    xfp_id,
    cenX = `Centroid X (µm)`,
    cenY = `Centroid Y (µm)`,
    cenZ = `Centroid Z (µm)`,
    vol = `Volume (µm3)`
  )

CFP_nuc_files <- list.files(dir, pattern = "CFP-nuclei", full.names = TRUE)
CFP_nuclei <- CFP_nuc_files %>% map_dfr(~read_csv(.)) %>%
  mutate (field = str_remove(`Item Name`," \\(registration corrected\\)")) %>%
  separate(field, into = c("sample","field"), sep = " ") %>%
  replace_na(list(field="1")) %>%
  mutate(xfp_id = paste("nucleus",str_remove(Name,"Object "))) %>%
  select(
    sample,
    #time,
    field,
    xfp_id,
    cenX = `Centroid X (µm)`,
    cenY = `Centroid Y (µm)`,
    cenZ = `Centroid Z (µm)`,
    vol = `Volume (µm3)`
  )

CFP_spot_files <- list.files(dir, pattern = "CFP-foci", full.names = TRUE)
CFP <- CFP_spot_files %>% map_dfr(~read_csv(.)) %>%
  mutate (field = str_remove(`Item Name`," \\(registration corrected\\)")) %>%
  separate(field, into = c("sample","field"), sep = " ") %>%
  replace_na(list(field="1")) %>%
  #mutate(xfp_id = paste("cfp",str_remove(Name,"Point "))) %>%
  mutate(xfp_id = paste("cfp",str_remove(Name,"Point ")),
        vol = NA) %>%
  select(
    sample,
    #time,
    field,
    xfp_id,
    cenX = `Position X (µm)`,
    cenY = `Position Y (µm)`,
    cenZ = `Position Z (µm)`,
    vol
  )

#full_data <- YFP
full_data <- bind_rows(YFP,RFP,CFP,CFP_nuclei)
#full_data <- bind_rows(YFP,RFP)
#full_data <- bind_rows(YFP,RFP,CFP_nuclei)
#full_data <- bind_rows(YFP,RFP,CFP)
micro_data <- full_data %>% select(-vol)
```

### Now split data into separate lists by sample (colony) and field

```{r}
data_list <- split(micro_data, ~micro_data$sample + micro_data$field,drop = TRUE) %>%
  map(select,-sample:-field) %>%
  map(remove_rownames) %>%
  map(column_to_rownames,var="xfp_id")

# distance matrix for each field (image)
dist_list <- map(data_list,dist)
# cluster info per field
clust_list <- map(dist_list,hclust,"complete") 
# cut at height=3 for individual nuclei
cluster_member_list <- map(clust_list,stats::cutree,h=3) # cutree masked by dendextend
```

### Plot dendrograms for every field using the nice plotting features of the dendextend package

```{r}
dend_list <- map(clust_list,as.dendrogram)
color_set <- c(rfp = "red", cfp = "blue",
               yfp = "yellow", nuc = "grey")
dend_list <- map(dend_list,set,"leaves_pch",19) # filled circles
dend_list2 <- map(dend_list,~set(.x,
                                 "leaves_col",
                   color_set[str_sub(labels(.x),1,3)]))
#iwalk(dend_list,~ plot(.x,main=.y))
#dend_list <- map(dend_list,highlight_branches_lwd,values = seq(5, 1, length.out = 1000))
#dend_list <- map(dend_list,color_branches,h=3,col=turbo(10)) # color by cluster group
#iwalk(dend_list,~ plot(.x,main=.y,ylim=c(0,5),horiz = FALSE)) #imap is 'indexed' so name is

saveRDS(dend_list2,"dend.RDS")
```

### Use cluster information to carve out distance info for those signals

#### separate out nuclear signal-containing clusters only

```{r}
nuclear_cluster_list <- map(cluster_member_list,function(x){
  y <- x[which(str_detect(names(x),"nucleus"))] # subsets cluster numbers named "nucleus..."
  return(x[x %in% y]) # subsets only cluster numbers that have nucleus
})
```


```{r}
#cluster_df <- imap(cluster_member_list,~ tibble(sample.field=.y,xfp_id=names(.x),cluster_number=.x)) %>%
cluster_df <- imap(nuclear_cluster_list,~ tibble(sample.field=.y,xfp_id=names(.x),cluster_number=.x)) %>%
  bind_rows() %>%
  separate(col=sample.field, into = c("sample","field"),sep = "\\.")

# can join volume info here maybe

#gives a large list of $sample.field.group : XFP IDs
cluster_sets <- cluster_df %>%
  split(~.$sample + .$field + .$cluster_number,drop = TRUE) %>%
 map(pull,xfp_id)

# make list of dist matrices by sample.field.group
#### package the code below into map function

#which(sapply(cluster_sets, function(e) is.element('nucleus', str_extract(e,"\\w+\\b"))))
#sapply(cluster_sets, function(e) is.element('nucleus', str_extract(e,"\\w+\\b")))

# predefine output data structure for `for` loop
cluster_set_dist <- vector("list", length(cluster_sets)) # this would not be necessary using map

names(cluster_set_dist) <- names(cluster_sets) # this would not be necessary using map

for(i in seq_along(cluster_sets)) {
  y <- names(cluster_sets[i])
  x <- cluster_sets[[i]]
  if (identical(x, character(0))) next # likely bc single fluor
  xfp <- x[which(str_detect(x,"fp \\d"))] # ignore CFP nuclei for distance matrices
  y_sub <- str_replace(y,"(.*)\\.\\w+","\\1") # drops after last "."
  k <- as.matrix(dist_list[[y_sub]]) # extracts colony.field distance list
  l <- k[rownames(k) %in% xfp, # subsets that distance matrix
         colnames(k) %in% xfp]
  l[!lower.tri(l)] <- NA # blanks upper tri
  cluster_set_dist[[i]] <- l
}

##################

### Make strings of fluorophore observations in df

fluor_signal_df <- cluster_df %>%
  arrange(xfp_id) %>%
  #filter(!str_detect(xfp_id,"nucleus")) %>%
  group_by(sample,field,cluster_number) %>%
  summarise(
    signals = reduce(xfp_id,paste,collapse = " "),
    signal_code = reduce(str_sub(xfp_id,1,3), paste, collapse = " ")
  ) %>%
  ungroup()
  
### 'position' of cluster is mean of all objects in x,y,z

cluster_pos <- cluster_df %>%
  left_join(micro_data) %>%
  group_by(sample,field,cluster_number) %>%
  summarise(
    loc.x = mean(cenX),
    loc.y = mean(cenY),
    loc.z = mean(cenZ)
  ) %>%
  ungroup()

#fluor_signal_df <- fluor_signal_df %>%



##################

### Get individual distances between each pair of fluors

### RY

RY <- map(cluster_set_dist,function(m) {

  y <- grep("yf",rownames(m),value = TRUE)
  r <- grep("rf",rownames(m),value = TRUE)
  c <- grep("cf",colnames(m),value = TRUE)
  
  # ry <- paste(r,y,sep = ".")
  # cy <- paste(c,y,sep = ".")
  # cr <- paste(c,r,sep = ".")
  
  if (length(y)+length(r)+length(c) > 3 |
      length(y)+length(r)+length(c) < 2) {
    rval <- c(NA,NA,NA)
    names(rval) <- c("ry","cy","cr")
    return(rval)
  } else {
    if(is_empty(r) | is_empty(y)) ry=NA
    else ry = min(m[r,y])
    if(is_empty(c) | is_empty(y)) cy=NA
    else cy = min(m[c,y])
    if(is_empty(c) | is_empty(r)) cr=NA
    else cr = min(m[c,r])
    rval <- c(ry,cy,cr)
    names(rval) <- c("ry","cy","cr")
    return(rval)
  }
  
    
})

#colnames = c("ry","cy","cr")

RYdf <- RY %>%
  as_tibble() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("sample.field.group") %>%
  rename(c("ry"=V1,"cy"=V2,"cr"=V3)) %>%
  separate(col=sample.field.group,into = c("sample","field","cluster_number"),sep = "\\.") %>%
  mutate(cluster_number = as.integer(cluster_number))


final_df <- fluor_signal_df %>%
  left_join(cluster_pos) %>%
  left_join(RYdf)

saveRDS(final_df,"cluster-meta.RDS")

table_one <- final_df %>%
  filter(str_detect(signals,"nucleus")) %>% # remove CPF spot only signals
  group_by(sample) %>%
  count(signal_code) %>%
  arrange(as.numeric(str_remove(sample,"c")),desc(n)) %>%
  ungroup()

table_two <- final_df %>%
  filter(signal_code == "cfp nuc rfp yfp") %>%
  mutate(coloc = case_when(
    (ry < cy) & (ry < cr) ~ "ry",
    (cy < ry) & (cy < cr) ~ "cy",
    (cr < cy) & (cr < ry) ~ "cr",
  )) %>%
  group_by(sample) %>%
  count(coloc) %>%
  arrange(as.numeric(str_remove(sample,"c")),desc(n))


signal_codes <- table_one %>%
  complete(sample,signal_code,fill = list(n=0)) %>%
  group_by(sample) %>%
  mutate(
    prop = n/sum(n)
  ) %>%
  ungroup() %>%
  select(-n) %>%
  pivot_wider(names_from = signal_code,values_from = prop) %>%
  column_to_rownames(var = "sample")

signal_dist <- dist(signal_codes)
signal_clust <- hclust(signal_dist)
signal_dend <- as.dendrogram(signal_clust) %>% set("labels_cex",0.5)

plot(signal_dend, horiz = TRUE)
```

Try:
  - dropping categories of signals that never get above a certain cutoff



Nice tables

```{r}
library(kableExtra)
kbl(table_one,
    col.names = c("Sample","Signals","n"),
    align = c('c','c','c'))

kbl(table_two,
    col.names = c("Sample "," co-localized ","n"),
    align = c('c','c','c'))


```



bring volume information


```{r}
vol_cluster <- full_data %>%
  left_join(cluster_df) %>%
  mutate(xfp = str_remove(xfp_id," \\d+"),.keep = "unused") %>%
  pivot_wider(id_cols = c(sample,field,cluster_number),
              names_from = xfp,
              values_from = c(cenX,cenY,cenZ,vol))



  ggplot(data.frame(unlist(vol_cluster$vol_nucleus)),aes(x=unlist.vol_cluster.vol_nucleus.)) +
  geom_histogram(bins = 100)
  
dens_yfp <- density(unlist(vol_cluster$vol_nucleus))
```




Plots

```{r}
final_df %>%
  filter(str_detect(signals,"nucleus")) %>% # remove CPF spot only signals
  select(ry,cy,cr) %>%
  pivot_longer(cols = everything(),names_to = "coloc_signal",values_to = "distance") %>%
  ggplot(aes(x=distance,color=coloc_signal)) +
  geom_density(size=1.5) +
  theme_bw()
```

```{r}
#as.dist(cluster_set_dist$c1.2.4) %>% hclust("complete") %>% plot(ylim=c(0,3))
dend <- as.dendrogram(hclust(as.dist(cluster_set_dist$c1.2.4),
                             "complete")) %>% 
  set("leaves_pch", 21) %>% 
  set("leaves_bg", "gold") %>%    
  set("leaves_cex", 2) %>% 
  set("leaves_col", "darkred")

plot(dend,horiz = TRUE,xlim=c(3,0))
```

Map cluster_set_dist to function that:
  as.dist
  hclust("complete")
  as.dendrogram
  
then save for use in app

2 or fewer nodes is a problem

```{r}
cluster_set_dend <- cluster_set_dist %>%
  map(function(x){
    as.dendrogram(hclust(as.dist(x),"complete"))
  })
```

